---
title: "Module 11"
author: "Patricia McGee"
date: '2022-04-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Instructions

Your assignment for module # 11 is to review Dr. Piwek posting on Tufte and  C. Minard in R  (Links to an external site.)and see if you can generate one of the visualization he discussed in his post that include:

i. Marginal histogram scatter plot
ii. Dot-dash plot in ggplot2
iii. Dot-dash plot in lattice
Or any one you like the best. 

## My Code

For this assignment, we are given the resource on Tufte. There are many graphical examples and here I will recreate a few of them.

```{r code, echo=TRUE}
# install.packages(c("CarletonStats", "devtools", "epanetReader", "fmsb", "ggplot2", "ggthemes", 
#                   "latticeExtra", "MASS", "PerformanceAnalytics", "psych", 
#                   "plyr", "prettyR", "plotrix", "proto", "RCurl", "reshape", "reshape2"))

library(CarletonStats)
library(devtools)
library(epanetReader)
library(fmsb)
library(ggplot2)
library(ggthemes)
library(latticeExtra)
library(MASS)
library(PerformanceAnalytics)
library(psych)
library(plyr)
library(prettyR)
library(plotrix)
library(proto)
library(RCurl)
library(reshape)
library(reshape2)

```

## Plots

Here are my plots.

```{r plots, echo=TRUE}

# Minimal Line Plot in Lattice
x <- 1967:1977
y <- c(0.5,1.8,4.6,5.3,5.3,5.7,5.4,5,5.5,6,5)
xyplot(y~x, xlab="", ylab="", pch=16, col=1, border = "transparent", type="o",
       abline=list(h = c(max(y),max(y)-1), lty = 2),
       scales=list(x=list(at=x,labels=x, fontfamily="serif", cex=1),
                   y=list(at=seq(1,6,1), fontfamily="serif", cex=1,
                          label=sprintf("$%s",seq(300,400,20)))),
       par.settings = list(axis.line = list(col = "transparent"), dot.line=list(lwd=0)),
       axis = function(side, line.col = "black", ...) {
         if(side %in% c("left","bottom")) {axis.default(side = side, line.col = "black", ...)}})
ltext(current.panel.limits()$xlim[2]/1.1, adj=1, fontfamily="serif", 
      current.panel.limits()$ylim[1]/1.3, cex=1,
      "Per capita\nbudget expandures\nin constant dollars")
ltext(current.panel.limits()$xlim[2]/1.1, adj=1, fontfamily="serif", 
      current.panel.limits()$ylim[1]/5.5, cex=1, "5%")

# Interactive minimal line chart with highcharter
library(highcharter)
x <- 1967:1977
y <- c(290,318,372,385,385,372,386,380,390,400,380)
d <- data.frame(x, y)
highchart() %>%
  hc_chart(type = "scatter") %>% 
  hc_subtitle(text = "Per capita budget expanditures in constant dollars") %>%
  hc_yAxis(labels = list(format = "${value}")) %>%
  hc_add_series(data = d) %>% 
  hc_add_theme(hc_theme_tufte()) %>%
  hc_exporting(enabled = TRUE, filename = "highchartgraph")

# Dot Dash Plot in ggplot2
ggplot(mtcars, aes(wt, mpg)) + geom_point() + geom_rug() + theme_tufte(ticks=F) + 
  xlab("Car weight (lb/1000)") + ylab("Miles per gallon of fuel") + 
  theme(axis.title.x = element_text(vjust=-0.5), axis.title.y = element_text(vjust=1))


```

## Marginal Histogram with Scatterplot

```{r marginal, echo=TRUE}
library(ggExtra)
p <- ggplot(faithful, aes(waiting, eruptions)) + geom_point() + theme_tufte(ticks=F)
ggMarginal(p, type = "histogram", fill="transparent")
```

