---
title: "Module 13"
author: "Patricia McGee"
date: '2022-04-12'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Assignment

Your assignment:

Create a simple animation and discussion the result in your blog posting. 
Reference to Yihui Xie's and animation R package website, click here. A note, his package now available via CRAM repository and not GitHub.  For Bryan's blog posting on Animation in R, click here 
A note: In order to post your work on blog, you need to save your animation.
The code for saving the image you created in R is as follow:
saveGIF({
    for (i in 1:10) plot(runif(10), ylim = 0:1)
})

## My Work

I am going to create an animated plot with the gapminder data. I will filter to include only data for the following countries: China, France, Germany, United Kingdom, and United States. I plan to show the change in population and change in life expectancy over time.


```{r code, echo=TRUE}
#install.packages("animation")
library(animation)
library(gapminder)
library(dplyr)
library(ggplot2)
library(gganimate)

mydata <- gapminder

mydata <- mydata %>%
  filter(country %in% c("China","France","Germany","United Kingdom","United States"))

```

## Creating our Animated Plot using animate Package

My first animation is using the animate package. Unfortunately, this package does not have clear instructions on how to set this up, so I have been unable to generate a graph other than the example provided by Dr. Friedman.


```{r animate, echo=TRUE}
saveGIF({
    for (i in 1:10) plot(runif(10), ylim = 0:1)
}, movie.name = "animationsample.gif")

```

## Creating an Animation using gganimate Package

For my second plot, I am going to use the gganimate package, which works with ggplot. This package is much easier to work with.

```{r gganimate, echo=TRUE}
ggplot(mydata, aes(year,lifeExp, color = country))+
  geom_line()+
  geom_point()+
  labs(title = "Country Life Expectancy by Year", x = "Year", y = "Life Expectancy")+
  transition_reveal(year)+
  ease_aes("linear")

anim_save("gganimateplot.gif")

```

