---
title: "LIS 4317 Final Project"
author: "Patricia McGee"
date: '2022-04-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Assignment



## My Information and Code

I have decided to use the Beer Profile and Ratings Dataset from Kaggle: https://www.kaggle.com/ruthgn/beer-profile-and-ratings-data-set

I will read in the csv file "beer_profile_and_ratings.csv" and save this as beers.


```{r data, echo=TRUE}
beers <- read.csv("beer_profile_and_ratings.csv")
head(beers)
summary(beers)

```

## Exploring and Cleaning the Data

First, we need to narrow this dataset down for our purposes. There are currently 3197 observations. 

We remove some fields that we don't need. We will remove Description and the Min.IBU and Max.IBU.

Then we will filter this to only select Styles. I have chosen 8 beer styles that I enjoy. They are: 
Blonde Ale - Belgian
Cream Ale
Lager - American
Pumpkin Beer
Red Ale - Irish
Stout - Irish Dry
Stout - Sweet / Milk
Wheat Beer - Hefeweizen

This brings us down to 248 observations. 


```{r cleaning, echo=TRUE}
library(dplyr)
beers <- beers[,-7:-8]
beers <- beers[,-5]

beers <- beers %>%
  rename(Name = ï..Name, FullBeerName = Beer.Name..Full.)

beers_filtered <- beers %>%
  filter(Style %in% c("Blonde Ale - Belgian","Cream Ale","Lager - American","Pumpkin Beer","Red Ale - Irish","Stout - Irish Dry","Stout - Sweet / Milk","Wheat Beer - Hefeweizen"))

top_beers <- beers_filtered %>%
  arrange(desc(review_overall))%>%
  top_n(n = 10, review_overall)

topABVbeers <- beers_filtered %>%
  arrange(desc(ABV))%>%
  top_n(n = 10, ABV)


```


## Plots

I chose to create plots showing the top 10 beers by overall rating and the top 10 beers by ABV. I also created a box plot showing the range of the overall rating by beer style for the 8 selected styles and a linear regression line showing the ABV for each beer style.

```{r plots, echo=TRUE}
library(ggplot2)
library(stringr)
library(forcats)

ggplot(beers_filtered, aes(Style, review_overall))+
  geom_boxplot()+
  coord_flip()+
  theme_minimal()+
  labs(title = "Overall Ratings of Specific Beer Styles", x = "Beer Style", 
       y = "Overall Rating: Scale of 1-5")

ggplot(beers_filtered, aes(review_overall, ABV))+
  geom_point(size = 1, alpha = 0.8, aes(color = Style))+
  geom_smooth(size = 0.6, aes(color = Style, fill = Style))+
  theme_minimal()+
  scale_color_brewer(palette = "Set2")+
  scale_fill_brewer(palette = "Set2")+
  facet_wrap(vars(Style))+
  labs(title = "Alcohol by Volume (ABV) and Overall Rating of Specific Beer Styles", 
       x = "Overall Rating: Scale of 1-5", y = "Alcohol by Volume (ABV) %")

top_beers %>%
  mutate(Beer = fct_reorder(Name,review_overall)) %>%
  ggplot(aes(review_overall, Beer))+
  geom_col(aes(fill = Style))+
  coord_cartesian(xlim = c(4.2,4.6))+
  scale_fill_brewer(palette = "Set2")+
  scale_y_discrete(labels = function(y) str_wrap(y, width = 30))+
  labs(title = "Top 10 Rated Beers of Selected Styles")+
  theme_minimal()

topABVbeers %>%
  mutate(Beer = fct_reorder(Name,ABV))%>%
  ggplot(aes(Beer, ABV))+
  geom_col(aes(fill = Style))+
  coord_flip()+
  scale_fill_brewer(palette = "Set2")+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 30))+
  labs(title = "Top 10 Beers by ABV with Beer Style")+
  theme_minimal()


```

