---
title: "LIS 4317 Final Project"
author: "Patricia McGee"
date: '2022-04-26'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Assignment

### The Project outline

Each student in the class will create his/her visual project. The student will share the link on Canvas. The grade for the final project will be scored based on your instructor and your classmates peer review comments and grades. 

### Objectives

The purpose of this project is to apply what you’ve learned in this class: design and statistical analysis apparatus. Your project needs to capture any type of data based on your interest and utilize plot.ly, Tableau or Rstudio. The software you picked has to be covered in class in order to tell a story of your data analysis you selected. You can employ any of the techniques and practices we outline in the class. From Time series analysis, part to whole, ranking, deviation analysis, distribution, correlation, multivariate, and even promising trends as discussed in Chapter 13 in our textbook.

#### Step 1. 

Begin searching for a dateset that captivates your interest. Your dataset should contain at least 50 observations and 5 to 20 variables.

#### Step 2.

You will provide a written documentation of your project objectives and accomplishments that will include a general research question and objectives. Those questions will contain your underlying theme of your project as outline below.

Short summary:

Problem description: Give a brief but precise description or definition of the problem or
hypothesis you set  to evaluate.
Related work: How does this problem and the method relate to the problems/methods have addressed in the Provide example of existing visual analytics that you liked.
Solution: How did you solve the problem? Describe the technical approach. Tell us what methodology you will take: Time series analysis, part to whole, ranking, deviation analysis, distribution, correlation, multivariate, and even promising trends.

#### Step 3.

Create a visualization of your data and discussion to your output. You can use any of the apps we covered in the class including: plot.ly, Tableau, RStudio and Python.   

## My Information and Code

I have decided to use the Beer Profile and Ratings Dataset from Kaggle: https://www.kaggle.com/ruthgn/beer-profile-and-ratings-data-set

I will read in the csv file "beer_profile_and_ratings.csv" and save this as beers.


```{r data, echo=TRUE}
beers <- read.csv("beer_profile_and_ratings.csv")
head(beers)
summary(beers)

```

## Exploring and Cleaning the Data

First, we need to narrow this dataset down for our purposes. There are currently 3197 observations. 

We remove some fields that we don't need. We will remove Description and the Min.IBU and Max.IBU.

Then we will filter this to only select Styles. I have chosen 8 beer styles that I enjoy. They are: 
Blonde Ale - Belgian
Cream Ale
Lager - American
Pumpkin Beer
Red Ale - Irish
Stout - Irish Dry
Stout - Sweet / Milk
Wheat Beer - Hefeweizen

This brings us down to 248 observations. 


```{r cleaning, echo=TRUE}
library(dplyr)
beers <- beers[,-7:-8]
beers <- beers[,-5]

beers <- beers %>%
  rename(Name = ï..Name, FullBeerName = Beer.Name..Full.)

beers_filtered <- beers %>%
  filter(Style %in% c("Blonde Ale - Belgian","Cream Ale","Lager - American",
                      "Pumpkin Beer","Red Ale - Irish","Stout - Irish Dry",
                      "Stout - Sweet / Milk","Wheat Beer - Hefeweizen"))

top_beers <- beers_filtered %>%
  arrange(desc(review_overall))%>%
  top_n(n = 10, review_overall)

topABVbeers <- beers_filtered %>%
  arrange(desc(ABV))%>%
  top_n(n = 10, ABV)


```


## Plots

I chose to create plots showing the top 10 beers by overall rating and the top 10 beers by ABV. I also created a box plot showing the range of the overall rating by beer style for the 8 selected styles and a linear regression line showing the ABV for each beer style.

```{r plots, echo=TRUE}
library(ggplot2)
library(stringr)
library(forcats)

beers_filtered %>%
  mutate(BeerStyle = fct_reorder(Style, review_overall))%>%
  ggplot(aes(BeerStyle, review_overall))+
  geom_boxplot()+
  coord_flip()+
  theme_minimal()+
  labs(title = "Overall Ratings of Specific Beer Styles", x = "Beer Style", 
       y = "Overall Rating: Scale of 1-5")

ggplot(beers_filtered, aes(review_overall, ABV))+
  geom_point(size = 1, alpha = 0.8, aes(color = Style))+
  geom_smooth(size = 0.6, aes(color = Style, fill = Style))+
  theme_minimal()+
  scale_color_brewer(palette = "Set2")+
  scale_fill_brewer(palette = "Set2")+
  facet_wrap(vars(Style))+
  labs(title = "Alcohol by Volume (ABV) and Overall Rating of Specific Beer Styles", 
       x = "Overall Rating: Scale of 1-5", y = "Alcohol by Volume (ABV) %")

top_beers %>%
  mutate(Beer = fct_reorder(FullBeerName,review_overall)) %>%
  ggplot(aes(review_overall, Beer))+
  geom_col(aes(fill = Style))+
  coord_cartesian(xlim = c(4.2,4.6))+
  scale_fill_brewer(palette = "Set2")+
  scale_y_discrete(labels = function(y) str_wrap(y, width = 30))+
  labs(title = "Top 10 Rated Beers of Selected Styles", y = "Beer Name", 
       x = "Overall Rating: Scale of 1-5")+
  theme_minimal()

topABVbeers %>%
  mutate(Beer = fct_reorder(FullBeerName,ABV))%>%
  ggplot(aes(Beer, ABV))+
  geom_col(aes(fill = Style))+
  coord_flip()+
  scale_fill_brewer(palette = "Set2")+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 30))+
  labs(title = "Top 10 Beers by ABV with Beer Style", x = "Beer Name", y = "% ABV")+
  theme_minimal()


```

